---
title: "STA304PS4JCodeAppendix2"
output: pdf_document
---

Preamble: This code begins by taking two survey datasets, from Nationscape and IPUMS, and organizing the questions and responses to these surveys into nicer groups that we can better analyze. Once these groups have been defined for each survey, their proportions have been displayed using a series of bar graphs. Next this code runs a regression model on the Nationscape survey data and we analyze both the validity and results of this model. Lastly, using post-stratification, we apply regression model to our IPUMS ACS Dataset to generate a forecast for the 2020 US Election, as well as the standard error on this forecast. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE}
# Load Packages, use install.package() if needed
library(haven)
library(tidyverse)
library(labelled)
library(jtools)
library(boot)
```

```{r}
# Read in the raw data (You might need to change this if you use a different dataset)
raw_data <- read_dta("ns20200625.dta")

# Transform into Factored form
raw_data <- labelled::to_factor(raw_data)
```


```{r}
# Select variable of interest
select_data <- raw_data %>%
  dplyr::select(vote_2020,
                age,
                gender,
                race_ethnicity,
                census_region,
                language
  )
```

```{r}
# omit NA Values and isolate data for voters of Trump or Biden
clean_data <- select_data %>%
  na.omit() %>%
  filter(vote_2020 %in% c("Joe Biden", "Donald Trump"))
```

```{r}
# grouping age into 5 categories, all other races besides "White" and "Black, or African American" into an "Other" category, all other languages besides "English" and "Spanish" into an "Other" category, and lastly remove all NA values from this data.
clean_data <- clean_data %>%
  mutate(age_group = cut(age, 
                         breaks = c(18, 30, 40, 50, 60, 100), 
                         right = FALSE,
                         labels = c("18 to 29", 
                                    "30 to 39", 
                                    "40 to 49", 
                                    "50 to 59",
                                    "60 +")
                         ),
         race_ethnicity = case_when(race_ethnicity == "Black, or African American" ~ "Black, or African American",
                                    race_ethnicity == "White" ~ "White",
                                    race_ethnicity != "Black, or African American" | race_ethnicity != "White" ~ "Other"),
         language = case_when(language == "Yes, we speak Spanish." ~ "Spanish",
                              language == "Yes, we speak a language other than Spanish or English." ~ "Other",
                              language == "No, we speak only English." ~ "English") %>%
           na.omit()

           
  )
```

# 2.2 Display of Survey Data

```{r, echo=FALSE}
# Graph of vote decisions of respondents 
clean_data %>% 
  ggplot(aes(x = vote_2020)) + 
  geom_bar(colour="black", fill = 'blue') + 
  labs(title = "Graph 1: Vote Decisions of Respondents", 
       x = "Vote Choice",
       y = "Number of Respondents"
       ) + 
  theme_minimal()

```



```{r, echo=FALSE}
# Bar graph of age proportions of respondents
clean_data %>% 
  ggplot(aes(x = age_group)) + 
  geom_bar(colour = "black", fill = "darkturquoise") +
  labs(title = "Graph 2: Age Proportions of Respondents", 
       x = "Age",
       y = "Number of Respondents"
       ) + 
  theme_minimal()
```

```{r, echo=FALSE}
# Bar graph of gender proportions of respondents 
clean_data %>% 
  ggplot(aes(x = gender)) + 
  geom_bar(colour = "black", fill = "firebrick2") + 
  labs(title = "Graph 3: Gender Proportions of Respondents", 
       x = "Gender",
       y = "Number of Respondents"
       ) +
  theme_minimal()

```

```{r, echo=FALSE}
# Bar graph of the Races of respondents
clean_data %>% 
  ggplot(aes(x = race_ethnicity)) + 
  geom_bar(colour = "black", fill = "orangered1") + 
  labs(title = "Graph 4: Races of Respondents", 
       x = "Race",
       y = "Number of Respondents"
       ) + 
  coord_flip() + 
  theme_minimal()
```

```{r, echo=FALSE}
# Bar graph of languages used by respondents 
clean_data %>% 
  ggplot(aes(x = language)) + 
  geom_bar(colour = "black", fill = "springgreen4") + 
  labs(title = "Graph 5: Languages Used by Respondents", 
       x = "Language Used at Home",
       y = "Number of Respondents"
       ) + 
  theme_minimal()

```

```{r, echo=FALSE}
# Bar graph of regional distribution of respondents
clean_data %>% 
  ggplot(aes(x = census_region)) + 
  geom_bar(colour = "black", fill = "mediumseagreen") + 
  labs(title = "Graph 6: Regional Distribution of Respondents", 
       x = "Vote Choice",
       y = "Number of Respondents"
       ) +
  theme_minimal()
```


# 2.3 ACS Data

```{r, include=FALSE}
raw_strat_data <- read_dta("usa_00002.dta")
```

```{r, include=FALSE}
raw_strat_data <- labelled::to_factor(raw_strat_data)
```

```{r, include=FALSE}
raw_select_strat_data <- raw_strat_data %>%
  select(age,
         sex,
         race,
         hispan,
         bpl,
         region,
         language,
         empstat
    
  )
```

```{r, include=FALSE}
strat_data <- raw_select_strat_data %>% 
  na.omit()
```

```{r, include=FALSE}
clean_strat_data <-
  strat_data %>%
  select(region, 
         sex, 
         age, 
         race, 
         bpl, 
         language) %>% 
  rename(gender = sex, 
         race_ethnicity = race, 
         census_region = region,
         foriegn_born = bpl)
```

```{r, include=FALSE}
clean_strat_data <- clean_strat_data %>%
  mutate(age_group = cut(as.numeric(age), 
                         breaks = c(18, 30, 40, 50, 60, 100), 
                         right = FALSE,
                         labels = c("18 to 29", 
                                    "30 to 39", 
                                    "40 to 49", 
                                    "50 to 59",
                                    "60 +")
                         ),
         census_region = case_when(census_region == "new england division" ~ "Northeast",
                                  census_region == "middle atlantic division" ~ "Northeast",
                                  census_region == "east north central div" ~ "Midwest",
                                  census_region == "west north central div" ~ "Midwest",
                                  census_region == "south atlantic division" ~ "South",
                                  census_region == "east south central div" ~ "South",
                                  census_region == "west south central div" ~ "South",
                                  census_region == "mountain division" ~ "West",
                                  census_region == "pacific division" ~ "West"
                                  ),
         language = case_when(language == "english" ~ "English",
                              language == "spanish" ~ "Spanish",
                              language != "english" | language != "spanish" ~ "Other"
                              ),
         race_ethnicity = case_when(race_ethnicity == "black/african american/negro" ~ "Black, or African American",
                                    race_ethnicity == "white" ~ "White",
                                    race_ethnicity != "black/african american/negro" | race_ethnicity != "white" ~ "Other"
                                    ),
         gender = case_when(gender == "male" ~ "Male",
                            gender == "female" ~ "Female")



         

    
  )
```

```{r, include=FALSE}
# Remove Data that doesn't fit the groups
clean_strat_data <- clean_strat_data %>%
  na.omit() %>%
  labelled::to_factor()
```

# 2.4 Display ACS Data

```{r, echo=FALSE}
# Bar graph of Age proportions of respondents for ACS data
clean_strat_data %>% 
  ggplot(aes(x = age_group)) + 
  geom_bar(colour = "black", fill = "darkturquoise") +
  labs(title = "Graph 7: Age Proportions of Respondents", 
       x = "Age",
       y = "Number of Respondents"
       ) +
  theme_minimal()
```

```{r, echo=FALSE}
# Bar graph for gender proportions of respondents for ACS data
clean_strat_data %>% 
  ggplot(aes(x = gender)) + 
  geom_bar(colour = "black", fill = "firebrick2") + 
  labs(title = "Graph 8: Gender Proportions of Respondents", 
       x = "Gender",
       y = "Number of Respondents"
       ) + 
  theme_minimal()
```

```{r, echo=FALSE}
# Bar graph for the races of respondents for ACS data
clean_strat_data %>% 
  ggplot(aes(x = race_ethnicity)) + 
  geom_bar(colour = "black", fill = "orangered1") + 
  labs(title = "Graph 9: Races of Respondents", 
       x = "Race",
       y = "Number of Respondents"
       ) + 
  coord_flip() + 
  theme_minimal()
```


```{r, echo=FALSE}
# Bar graph for the languages used by respondents for ACS data
clean_strat_data %>% 
  ggplot(aes(x = language)) + 
  geom_bar(colour = "black", fill = "springgreen4") + 
  labs(title = "Graph 10: Languages Used by Respondents", 
       x = "Language Used at Home",
       y = "Number of Respondents"
       ) + 
  theme_minimal()

```

```{r, echo=FALSE}
# Bar graph for the regional distribution of respondents for ACS data
clean_strat_data %>% 
  ggplot(aes(x = census_region)) + 
  geom_bar(colour = "black", fill = "mediumseagreen") + 
  labs(title = "Graph 11: Regional Distribution of Respondents", 
       x = "Vote Choice",
       y = "Number of Respondents"
       ) + 
  theme_minimal()
```


## 3 Model Discussion and Development 

# 3.1 Regression Model

```{r, include=FALSE}
# Initial regression model
survey_glm <- glm(as.numeric(vote_2020 == "Donald Trump") ~ 
                    age_group + 
                    gender +
                    census_region + 
                    race_ethnicity + 
                    language,
                  family = 'binomial',
                  data = clean_data
)
```

```{r, include = FALSE}
# Regression model used to plot coefficients
survey_glm_plot <- glm(vote_2020 ~ 
                    age_group + 
                    gender +
                    census_region + 
                    race_ethnicity + 
                    language,
                  family = 'binomial',
                  data = clean_data
)
```

# 3.2 Model Validation

```{r, include = FALSE}
# Calculating the 10 fold cross validation for model 
cv.glm(clean_data, survey_glm, K = 10)
```

```{r, include = FALSE}
# create 2 data sets to test Msres and MSPE
# adding id number for respondents in data sets 

test_data_MSPE <- clean_data %>%
  mutate(id = row_number())

test_data_msres <- clean_data %>% 
  mutate(id = row_number())
```

```{r, warning = FALSE, include = FALSE}
# Splicing Data into Testing and Training Set
set.seed(5799893)
glm_test_MSPE <- sample_n(test_data_MSPE, 1295)
glm_test_msres <- test_data_msres %>%
  anti_join(test_data_MSPE)
```

```{r, include = FALSE}
#Model validation, comparing MSPE and MSres
# fitting sample dataset to our model
glm_test1 <- glm(as.numeric(vote_2020 == "Donald Trump") ~ 
                    age_group + 
                    gender +
                    census_region + 
                    race_ethnicity + 
                    language,
                  family = 'binomial',
                  data = test_data_msres)

glm_test2 <- glm(as.numeric(vote_2020 == "Donald Trump") ~ 
                    age_group + 
                    gender +
                    census_region + 
                    race_ethnicity + 
                    language,
                  family = 'binomial',
                  data = test_data_MSPE)
                      
#calculation of the mspe for sample dataset
mspe = sum(resid(glm_test2)^2)/ length(test_data_MSPE$vote_2020)
#calculation of the MSres for our main dataset
# 13 represent the degrees of freedom in our model 
msres = sum(resid(glm_test1)^2)/(length(test_data_msres$vote_2020) - 13)
mspe
msres
```



```{r, echo  FALSE}
# Display summary statistics of our final model
broom::tidy(survey_glm) %>% 
  kable(digits = 2)
```



# 4 results

# Regression
```{r, echo = FALSE}
# Display predictor coefficients with distribution error and confidence interval 

plot_summs(survey_glm_plot, plot.distributions = TRUE, inner_ci_level = .95)
```

```{r, echo = FALSE}
# Display Cook’s distance for observations 
plot(survey_glm, which = 4, id.n = 3)
```

# Forecast

```{r, include = FALSE}
# Create Cells for our ACS Data
cell_counts <- clean_strat_data %>% 
  group_by(age_group, gender, census_region, race_ethnicity, language) %>%
  count() %>%
  mutate(proportion = n/2603150
         )

```

```{r,include=FALSE}
# Create Vector including all Prediction Values and standard errors
predictions <- predict.glm(survey_glm, se.fit = TRUE, cell_counts, type = "response")
```


```{r, include=FALSE}
# Add Predictions and standard errors to cell counts
cell_counts2 <- cbind(cell_counts, predictions$fit, predictions$se.fit)
```

```{r, include=FALSE}
# First Forecasting Result
vote_pred <- cell_counts2$proportion * cell_counts2$...8
sum(vote_pred)
```

```{r, include=FALSE}
# Calculating Lower and Upper Predictions based on Standard Error for each individual cell
lower_pred = cell_counts2$...8 - cell_counts2$...9
upper_pred = cell_counts2$...8 + cell_counts2$...9
cell_counts2 <- cbind(cell_counts2, lower_pred, upper_pred)
```

```{r, include=FALSE}
#  Rename the new Columns from Forecasting
cell_counts2 <- cell_counts2 %>% rename(
  prediction = ...8,
  StdError = ...9,
  LowerBound = ...10,
  UpperBound = ...11
)
```

```{r, include = FALSE}
# Calculate total Standard Errors 
lower_vote_pred <- cell_counts2$proportion * cell_counts2$LowerBound
sum(lower_vote_pred)
upper_vote_pred <- cell_counts2$proportion * cell_counts2$UpperBound
sum(upper_vote_pred)
```


















